---
title: "R Script: Does fluctuating selection maintain variation in nest defense behavior in Arctic peregrine falcons (Falco peregrinus tundrius)?"
output: github_document
html_document:
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r LoadData, include=T, warning=F, message=F, results='hide'}
#required packages
require(readr)
require(lme4)
require(MCMCglmm)
require(dplyr)
require(ggplot2)
require(rlang)
require(tidyverse)
require(rlang)
require(ggeffects)
require(arm)
require(bayestestR)

# Set wd
setwd("C:/Users/nickg/OneDrive/Desktop/R projects/krmp_nest-defense")
#load data
data<-read.csv("C:/Users/nickg/OneDrive/Desktop/R projects/krmp_nest-defense/data/nest_defense_test_final.csv")
names(data)
```

##### **Make Year,ID, and ID_Series factors. Inspect raw data and transform.**

```{r factor, include=T, warning=F, message=F,attr.source='.numberLines'}

data$Year2<-as.factor(data$Year2)
data$ID_Series<-as.factor(data$ID_Series)
data$ID<-as.factor(data$ID)

summary(data$MinDisRaw)
data$MinDis100<-ifelse(data$MinDisRaw>100,100,data$MinDisRaw)
data$LogMin<-log(data$MinDis100+1)
```

##### **Minimum distance is capped at 100 meters (i.e., >100 is not a response)**

```{r subset, include=T, warning=F, message=F, fig.height=5, fig.width=5}

data2<-subset(data, data$MinDisRaw<=100)
hist(data2$MinDisRaw)


data2$logMin<-log(data2$MinDisRaw+1)
hist(data2$logMin)

```

### **1. Sources of variation in nest defense**

- *Univariate analyses of each of the three measures to i) evaluate if all three behaviors are representations of nest defense, and ii) choose best measure to use in main text.*





#### **Univariate model-- Log-transformed minimum distance**

```{r mod1, include=T, warning=F, message=F,fig.height=5, fig.width=5}
#Minimum distance---- 
require(lme4)
m1<-lmer(logMin~ Sex + NestStage + Year2+ (1|ID)+(1|Observer) ,
         data=data2) 

#summary
summary(m1)
#model check
plot(m1)
hist(resid(m1)) # residuals
lattice::qqmath(m1) #normality of errors
```


##### **Simulate posterior distribution--Log-transformed minimum distance**
```{r, simM1, include=T, warning=F, message=F}
sm1<-sim(m1)
smfixef=sm1@fixef
smranef=sm1@ranef
smfixef=as.mcmc(smfixef)
posterior.mode(smfixef)
HPDinterval(smfixef)


bID<-sm1@ranef$ID
bvar<-as.vector(apply(bID, 1, var)) ##between individual variance posterior distribution
bvar<-as.mcmc(bvar)
posterior.mode(bvar )## mode of the distribution
HPDinterval(bvar)

bObs<-sm1@ranef$Observer
ObsVar<-as.vector(apply(bObs, 1, var)) ##between individual variance posterior distribution
ObsVar<-as.mcmc(ObsVar)
posterior.mode(ObsVar )## mode of the distribution
HPDinterval(ObsVar)

rvar<-sm1@sigma^2
rvar<-as.mcmc(rvar)
posterior.mode(rvar)
HPDinterval(rvar)
```

##### **Repeatability--Log-transformed minimum distance**

```{r, rptM1, include=T, warning=F, message=F}
r <- bvar/(bvar + rvar) ###individual
posterior.mode(r)
HPDinterval(r)
```

##### **Univariate model-- Log-transformed truncated minimum distance to the observer (i.e., >100m is not response)**

```{r, mod2, include=T, warning=F, message=T,fig.height=5, fig.width=5}
###min distance (truncated to max of 100)----
data$MinDis100<-ifelse(data$MinDisRaw>100,100,data$MinDisRaw)
data$logMinDis100<-log(data$MinDis100+1)
hist(data$logMinDis100)
m1.1<-lmer(logMinDis100~ Sex + NestStage + Year2+ (1|ID)+(1|Observer) ,
         data=data) 
#summary
summary(m1.1)
#model check
plot(m1.1)
hist(resid(m1.1)) # residuals
lattice::qqmath(m1.1) #normality of errors

```

##### **Simulate posterior distribution--Log-transformed truncated minimum distance to the observer (i.e., >100m is not response)**

```{r, simM2, include=T, warning=F, message=F}
sm1<-sim(m1.1)
smfixef=sm1@fixef
smranef=sm1@ranef
smfixef=as.mcmc(smfixef)
posterior.mode(smfixef)
HPDinterval(smfixef)

bID<-sm1@ranef$ID
bvar<-as.vector(apply(bID, 1, var)) ##between individual variance posterior distribution
bvar<-as.mcmc(bvar)
posterior.mode(bvar )## mode of the distribution
HPDinterval(bvar)

bObs<-sm1@ranef$Observer
ObsVar<-as.vector(apply(bObs, 1, var)) ##between individual variance posterior distribution
ObsVar<-as.mcmc(ObsVar)
posterior.mode(ObsVar )## mode of the distribution
HPDinterval(ObsVar)

rvar<-sm1@sigma^2
rvar<-as.mcmc(rvar)
posterior.mode(rvar)
HPDinterval(rvar)

```

##### **Repeatability--Log-transformed truncated minimum distance to the observer (i.e., >100m is not response)**

```{r, rptM2, include=T, warning=F, message=F}
r <- bvar/(bvar + rvar)        ###individual
posterior.mode(r)
HPDinterval(r)
```

#### **Univariate model-- Number of Stoops**

```{r, mod3, include=T, warning=F, message=F, fig.height=5, fig.width=5}

###Number of stoops----
#names(data)
m2<-glmer(Dives~ Sex + NestStage + Year2+ (1|ID)+ (1|Observer) ,
         data=data, family="poisson") 
#summary
summary(m2)
```

##### **Simulate posterior distribution--Number of Stoops**

```{r, m2post, include=TRUE, warning=F, message=F}

sm2<-sim(m2)
smfixef=sm2@fixef
smranef=sm2@ranef
smfixef=as.mcmc(smfixef)
posterior.mode(smfixef)
HPDinterval(smfixef)

bID<-sm2@ranef$ID
bvar<-as.vector(apply(bID, 1, var)) ##between individual variance posterior distribution
bvar<-as.mcmc(bvar)
posterior.mode(bvar )## mode of the distribution
HPDinterval(bvar)

bObs<-sm2@ranef$Observer
ObsVar<-as.vector(apply(bObs, 1, var)) ##between individual variance posterior distribution
ObsVar<-as.mcmc(ObsVar)
posterior.mode(ObsVar )## mode of the distribution
HPDinterval(ObsVar)
```

##### **Repeatability-- Number of Stoops**

```{r, rptm3, include=T, warning=F, message=F}

r <- bvar/(bvar + 1)        ###individual
posterior.mode(r)
HPDinterval(r)

```

#### **Univariate model-- Log-transformed FID**

```{r, mod4, include=T, warning=F, message=F, fig.height=5, fig.width=5}
###FID----
data3<-subset(data, data$FID>0)
hist(data3$FID)
data3$logFID<-log(data3$FID+1)
hist(data3$logFID)

m3<-lmer(logFID~ Sex + NestStage + Year2+ (1|ID)+(1|Observer) ,
         data=data3) 
#summary
summary(m3)
#model check
plot(m3)
hist(resid(m3)) # residuals
lattice::qqmath(m3) #normality of errors
```

##### **Simulate posterior distribution-- Log-transformed FID**

``` {r, simM3, include=T, warning=F, message=F}
sm3<-sim(m3)
smfixef=sm3@fixef
smranef=sm3@ranef
smfixef=as.mcmc(smfixef)
posterior.mode(smfixef)
HPDinterval(smfixef)

bID<-sm3@ranef$ID
bvar<-as.vector(apply(bID, 1, var)) ##between individual variance posterior distribution
bvar<-as.mcmc(bvar)
posterior.mode(bvar )## mode of the distribution
HPDinterval(bvar)

bObs<-sm3@ranef$Observer
ObsVar<-as.vector(apply(bObs, 1, var)) ##between individual variance posterior distribution
ObsVar<-as.mcmc(ObsVar)
posterior.mode(ObsVar )## mode of the distribution
HPDinterval(ObsVar)

rvar<-sm1@sigma^2
rvar<-as.mcmc(rvar)
posterior.mode(rvar)
HPDinterval(rvar)
```

##### **Repeatability-- Log-transformed FID**

```{r, rptM3, include=T, warning=F, message=F}
r <- bvar/(bvar + rvar)        ###individual
posterior.mode(r)
HPDinterval(r)
```


### **2. Multivariate model with 9 nest defense traits (across contexts)**

- *Since we had variable effects of nest stage with each behavior measured, we constructed three  multivariate models to understand the correlation of each measure of nest defense with nest stage.*

```{r, load, include=T, warning=F, message=F, results='hide', echo=F}
require(readr)
# Set wd
setwd("C:/Users/nickg/OneDrive/Desktop/R projects/krmp_nest-defense")
data3c<- read_csv("data/nest_defense_test_final_20220318.csv")
names(data3c)

data_3c<-subset(data3c, data3c$FID>0)


data_3c$logLFID<-log(data_3c$LFID+1) 
data_3c$logIFID<-log(data_3c$IFID+1)
data_3c$logPFID<-log(data_3c$PFID+1)

data_3c$logLMinDis<-log(data_3c$LMinDis+1)
data_3c$logIMinDis<-log(data_3c$IMinDis+1)
data_3c$logPMinDis<-log(data_3c$PMinDis+1)


data_3c<-as.data.frame(data_3c)
```


#### **Multivariate: FID across breeding contexts**

```{r, mvFID, include=T, warning=F, message=F,fig.height=8, fig.width=12}
prior3var=list(R=list(V=diag(3),nu=3.002),G=list(G1=list(V=diag(3), nu=3.002))) #3 trait prior w/one random effect

m3FID<-MCMCglmm(cbind(logLFID, logIFID,logPFID)~(trait-1),
             random=~us(trait):ID,
             rcov=~idh(trait):units,
             family=c("gaussian","gaussian", "gaussian"),
             prior=prior3var,
             nitt=1300000,thin=1000,burnin=300000, 
             data=data_3c,verbose=F)

summary(m3FID)
plot(m3FID)

```

##### **Among-individual covariance-FID**

```{r, amongCov88, include=T, warning=F}

c1 <- posterior.cor(m3FID$VCV[,1:9])
round(apply(c1,2,mean),2)
round(apply(c1,2, quantile, c(0.025, 0.975)),2)

```


#### **Multivariate: minimum approach distance across breeding contexts**

```{r, mvMin, include=T, warning=F, message=F,fig.height=8, fig.width=12}
m3Min<-MCMCglmm(cbind(logLMinDis, logIMinDis,logPMinDis)~(trait-1),
                random=~us(trait):ID,
                rcov=~idh(trait):units,
                family=c("gaussian","gaussian", "gaussian"),
                prior=prior3var,
                nitt=1300000,thin=1000,burnin=300000, 
                data=data_3c,verbose=FALSE)

summary(m3Min)
plot(m3Min)
```

##### **Among-individual covariance--minimum approach distance across breeding contexts**

```{r, amongCoMin, include=T, warning=F}

c1 <- posterior.cor(m3Min$VCV[,1:9])
round(apply(c1,2,mean),2)
round(apply(c1,2, quantile, c(0.025, 0.975)),2)
```

#### **Multivariate: number of dives across breeding contexts**

```{r, mvStoop, include=T, warning=F, message=F,fig.height=8, fig.width=12}
m3Dives<-MCMCglmm(cbind(Ldives, Idives,Pdives)~(trait-1),
                random=~us(trait):ID,
                rcov=~idh(trait):units,
                family=c("poisson","poisson", "poisson"), 
                prior=prior3var,
                nitt=1300000,thin=1000,burnin=300000, 
                data=data_3c,verbose=FALSE)

summary(m3Dives)
plot(m3Dives)
```

##### **Among-individual covariance--number of dives across breeding contexts**

```{r, amongCoStoop, include=T, warning=F, message=F}

c1 <- posterior.cor(m3Dives$VCV[,1:9])
round(apply(c1,2,mean),2)
round(apply(c1,2, quantile, c(0.025, 0.975)),2)

```


### **3. Bivariate model with minimum distance and number of dives**

- *Evaluate covariance between each measure of nest defense*

``` {r, 3bivariate, include=T, warning=F, message=F,fig.height=8, fig.width=12}
prior_2var= list(R = list(V = diag(2), nu = 1.002),
                 G = list(G1 = list(V = diag(2), nu = 2,
                                    alpha.mu = rep(0,2),
                                    alpha.V = diag(25^2,2)))) # 3-trait prior w/one random effect
#names(data)

model.2var <- MCMCglmm(
  cbind(LogMin, Dives) ~ trait - 1,
  random = ~ us(trait):ID,
  rcov = ~ us(trait):units,
  family = c("gaussian", "poisson"),
  data = data,
  prior = prior_2var,
  verbose = FALSE,
  nitt = 1300000, thin = 1000, burnin = 300000
  )


summary(model.2var)
plot(model.2var)
```

#### **Among-individual and within-individual covariance--minimum distance and number of dives**

```{r, amongCoBivar, include=T, warning=F, message=F}

c1<- posterior.cor(model.2var$VCV[,1:4])
round(apply(c1,2,mean),2)
round(apply(c1,2, quantile, c(0.025, 0.975)),2)

#within individual covariance
c2<- posterior.cor(model.2var$VCV[,5:8])
round(apply(c2,2,mean),2)
round(apply(c2,2, quantile, c(0.025, 0.975)),2)
```


### **4. Univariate model to estimate short-versus long-term repeatability with interaction**
- *Interaction is included since we predicted that females would be more defensive with increasing nest stages.*

```{r, univariateSvLr, include= T, warning= F, message=F}

#no sex:nest stage interaction... Table S3
require(lme4)
m1.0<-lmer(LogMin~ (Sex + NestStage)^2 + Year2+ (1|ID) + (1|ID_Series) ,
         data=data) 
summary(m1.0)
anova(m1.0)
```

##### **Simulate posterior distribution**

```{r, simm1.0, include=T, warning=F, message=F}

smod<-sim(m1.0,1000)
posterior.mode(as.mcmc(smod@fixef))
HPDinterval(as.mcmc(smod@fixef))
##Between individual variance
bID<-smod@ranef$ID[,,1]
bvar<-as.vector(apply(bID, 1, var)) ##ID variance posterior distribution
require(MCMCglmm)
bvar<-as.mcmc(bvar)
posterior.mode(bvar )## mode of the distribution
HPDinterval(bvar)
##Between individual variance, ID_Series
bID2<-smod@ranef$ID_Series[,,1]
bvar2<-as.vector(apply(bID2, 1, var)) ##ID_Series variance posterior distribution
require(MCMCglmm)
bvar2<-as.mcmc(bvar2)
posterior.mode(bvar2 )## mode of the distribution
HPDinterval(bvar2)
###residual variance
rvar<-smod@sigma^2
rvar<-as.mcmc(rvar)
posterior.mode(rvar)
HPDinterval(rvar)

```

##### **Long-term and short-term repeatability**

```{r, rptLvS, include=T, warning=F, message=T}
### Long-term repeatability
r<-bvar/(rvar+bvar +bvar2 )
posterior.mode(r)
HPDinterval(r)  ##repeatability 
###Short-term
r1<-(bvar +bvar2  )/(rvar+bvar +bvar2 )
posterior.mode(r1)
HPDinterval(r1)  ##repeatability

```

### **Main effects similar without including interactions.**

- *We dropped the interaction since it was not significant. Results for main text Table 1*

``` {r, unvME, include=T, warning=F, message=F, fig.height=5, fig.width=5}

cor.test(data$Stage, data$NumberOfChicks)#correlation b/n nest stage and number of chicks. r=0.78 thus drop Number of Chicks from model 
m1<-lmer(LogMin~ Sex + NestStage +Year2+ (1|ID) + (1|ID_Series) ,
         data=data) 
#summary
summary(m1)
#model check
plot(m1)
hist(resid(m1)) # residuals
lattice::qqmath(m1) #normality of errors
```

##### **Simulate posterior distribution-- Univariate model with no interaction**

```{r, unMESim, include=T, warning=F, message=T}
#simulated parameters
require(dplyr)
library(arm)
require(MCMCglmm)
smod<-sim(m1,1000)
posterior.mode(as.mcmc(smod@fixef))
HPDinterval(as.mcmc(smod@fixef))
##Between individual variance
bID<-smod@ranef$ID[,,1]
bvar<-as.vector(apply(bID, 1, var)) ##ID variance posterior distribution
require(MCMCglmm)
bvar<-as.mcmc(bvar)
posterior.mode(bvar )## mode of the distribution
HPDinterval(bvar)
##Between individual variance, ID_Series
bID2<-smod@ranef$ID_Series[,,1]
bvar2<-as.vector(apply(bID2, 1, var)) ##ID_Series variance posterior distribution
require(MCMCglmm)
bvar2<-as.mcmc(bvar2)
posterior.mode(bvar2 )## mode of the distribution
HPDinterval(bvar2)
###residual variance
rvar<-smod@sigma^2
rvar<-as.mcmc(rvar)
posterior.mode(rvar)
HPDinterval(rvar)

```

#### **Contrasts differences among-sexes**

```{r, contrast, include=T, warning=F, message=T}
#compute difference among-sexes

p222<-smod@fixef[,1:2]#gather posterior for intercept(female) and Male 
p222<-as.data.frame(p222)#make a dataframe to manipulate easier
colnames(p222)<-c("Female", "SexMale")#change column names
p22<-p222%>%
  summarise(Male= Female+ SexMale)#add posteriors for Intercept to each value of Male

con<-as.data.frame(c(p222, p22))#combine estimates into one dataframe

difference<-con%>%
  summarise(diff= Female - Male)#compute difference

contrast<-as.data.frame(c(con, difference))#dataframe for contrasts

require(bayestestR)
describe_posterior(contrast)# pd is 87% for difference among sexes 
contable<-describe_posterior(contrast)#estimates for difference b/n male and female nest defense

contab<-contable%>%filter(Parameter=="diff")


plot(pd(contrast$diff))+theme_classic()

ggplot(contrast, aes(x=diff)) + geom_density(fill="orange") +
  geom_vline(xintercept = median(contrast$diff), color="red", size=1)+
   geom_vline(xintercept = (contab$CI_low), color="black", size=1, linetype="longdash")+
  geom_vline(xintercept = (contab$CI_high), color="black", size=1, linetype="longdash")+
  labs(x= "Difference", y="Density")+
  ggtitle(label ="Difference between male and female defense",
          subtitle = "Posterior Distribution, Median, and Highest Density Interval")+
  theme(plot.title = element_text(face = "bold", hjust=0.5), plot.subtitle=element_text(hjust=0.5))
```

#### **Contrast differences among-nest stages**

```{r, contrast2, include=T, warning=F, message=F}


p2223<-smod@fixef[,3:4]#gather posterior for incubation and provisioning nest stages 
p2223<-as.data.frame(p2223)#make a dataframe to manipulate easier
p2224<-smod@fixef[,1]#extract intercept (estimate for egg-laying)
p2224<-as.data.frame(p2224)
colnames(p2224)<-"Intercept"

nestcon<-as.data.frame(c(p2224, p2223))

#compute estimates for each nest stage
neststage<-nestcon%>%
  summarise(Incubation   = Intercept  + NestStage2Incubating,
            Provisioning = Intercept + NestStage3Provisioning,
            EggLaying    = Intercept)#add posteriors for Intercept to each value of Male

#compute differences between nest stage groups
contrasts<-neststage%>%
  summarise(diff_ie = Incubation-EggLaying,
            diff_ip = Incubation-Provisioning,
            diff_ep = EggLaying-Provisioning)
require(bayestestR)
describe_posterior(contrasts)#median, 95% CrI and PD for each difference

contable<-describe_posterior(contrasts)#estimates for difference b/n male and female nest defense
diff_ie<-contable%>%filter(Parameter=="diff_ie")
diff_ip<-contable%>%filter(Parameter=="diff_ip")
diff_ep<-contable%>%filter(Parameter=="diff_ep")

# incubation vs egg-laying
p3<-ggplot(contrasts, aes(x=diff_ie)) + geom_density(fill="orange") +
  geom_vline(xintercept = median(contrasts$diff_ie), color="red", size=1)+
   geom_vline(xintercept = (diff_ie$CI_low), color="black", size=1, linetype="longdash")+
  geom_vline(xintercept = (diff_ie$CI_high), color="black", size=1, linetype="longdash")+
  labs(x= "Difference", y="Density")+
  ggtitle(label ="Difference between incubation and egg laying",
          subtitle = "Posterior Distribution, Median, and Highest Density Interval")+
  theme(plot.title = element_text(face = "bold", hjust=0.5), plot.subtitle=element_text(hjust=0.5))
p3
# pd plot incubation vs egg-laying
plot(pd(contrasts$diff_ie))+theme_classic()


# incubation vs provisioning
p2<-ggplot(contrasts, aes(x=diff_ip)) + geom_density(fill="orange") +
  geom_vline(xintercept = median(contrasts$diff_ip), color="red", size=1)+
   geom_vline(xintercept = (diff_ip$CI_low), color="black", size=1, linetype="longdash")+
  geom_vline(xintercept = (diff_ip$CI_high), color="black", size=1, linetype="longdash")+
  labs(x= "Difference", y="Density")+
  ggtitle(label ="Difference between incubation and provisioning",
          subtitle = "Posterior Distribution, Median, and Highest Density Interval")+
  theme(plot.title = element_text(face = "bold", hjust=0.5), plot.subtitle=element_text(hjust=0.5))
p2
# pd plot incubation vs egg-laying
plot(pd(contrasts$diff_ip))+theme_classic()


# egg laying vs provisioning
p1<-ggplot(contrasts, aes(x=diff_ep)) + geom_density(fill="orange") +
  geom_vline(xintercept = median(contrasts$diff_ep), color="red", size=1)+
   geom_vline(xintercept = (diff_ep$CI_low), color="black", size=1, linetype="longdash")+
  geom_vline(xintercept = (diff_ep$CI_high), color="black", size=1, linetype="longdash")+
  labs(x= "Difference", y="Density")+
  ggtitle(label ="Difference between incubation and provisioning",
          subtitle = "Posterior Distribution, Median, and Highest Density Interval")+
  theme(plot.title = element_text(face = "bold", hjust=0.5), plot.subtitle=element_text(hjust=0.5))
p1
# pd plot egg laying vs provisioning
plot(pd(contrasts$diff_ep))+theme_classic()

```

### **Contrast difference among years**

```{r, contrast5, include=T, warning=F, message=T}


p55<-smod@fixef[,5]#gather posterior for intercept(female) and Male 
p55<-as.data.frame(p222)#make a dataframe to manipulate easier
colnames(p55)<-c("Year2019")#change column names
p2224<-smod@fixef[,1]#extract intercept (estimate for egg-laying)
p2224<-as.data.frame(p2224)
colnames(p2224)<-"Intercept"

yearcon<-as.data.frame(c(p2224, p55))

p2299<-yearcon%>%
  summarise(Y2019 = Intercept+ Year2019)#add posteriors for Intercept(2018) to each value of Male

ycon<-as.data.frame(c(yearcon, p2299)) 


difference_y<-ycon%>%
  summarise(diff_year= Intercept - Y2019)#compute difference


require(bayestestR)
describe_posterior(difference_y)# pd is 87% for difference among sexes 
contable<-describe_posterior(difference_y)#estimates for difference b/n male and female nest defense




ggplot(difference_y, aes(x=diff_year)) + geom_density(fill="orange") +
  geom_vline(xintercept = median(difference_y$diff_year), color="red", size=1)+
   geom_vline(xintercept = (contable$CI_low), color="black", size=1, linetype="longdash")+
  geom_vline(xintercept = (contable$CI_high), color="black", size=1, linetype="longdash")+
  labs(x= "Difference", y="Density")+
  ggtitle(label ="Difference between 2018 and 2019",
          subtitle = "Posterior Distribution, Median, and Highest Density Interval")+
  theme(plot.title = element_text(face = "bold", hjust=0.5), plot.subtitle=element_text(hjust=0.5))

plot(pd(difference_y$diff_year))+theme_classic()
```

##### **Long-term and short-term repeatability--Univariate model with no interaction**

``` {r, lvsrpt, include=T, warning=F, message=F}
### Long-term repeatability
r<-bvar/(rvar+bvar +bvar2 )
posterior.mode(r)
HPDinterval(r)  ##repeatability 
###Short-term
r1<-(bvar +bvar2  )/(rvar+bvar +bvar2 )
posterior.mode(r1)
HPDinterval(r1)  ##repeatability
```


##### **Plot--Univariate model with no interaction**

``` {r, plot1, include=T, warning=F, message=F,fig.width=10, fig.height=5}
#load packages for predictive plot
require(ggplot2)
require(ggeffects)
require(cowplot)

d18<-subset(data, data$Year=="2018")
d19<-subset(data, data$Year=="2019")

#plot
p18<-ggplot(d18, aes(x = NestStage, y = MinDisRaw, factor=NestStage, fill=Sex)) + geom_boxplot() +
  scale_y_continuous("Minimum approach distnace to observer (m)",limits = c(0, 100)) +
  scale_x_discrete("Nest Stage", labels=c('Egg-laying', 'Incubating', 'Provisioning'))+ 
  guides(fill="none")+
  guides(color="none")+ theme_classic(base_size = 12)  +
  ggtitle(label ="2018",
          subtitle = "")+
  theme(plot.title = element_text(face = "bold", hjust=0.5), plot.subtitle=element_text(hjust=0.5))



p19<-ggplot(d19, aes(x = NestStage, y = MinDisRaw, factor=NestStage, fill=Sex)) + geom_boxplot() +
  scale_y_continuous("Minimum approach distnace to observer (m)",limits = c(0, 100)) +
  scale_x_discrete("Nest Stage", labels=c('Egg-laying', 'Incubating', 'Provisioning'))+ 
  guides(color="none")+ theme_classic(base_size = 12)  +
  ggtitle(label ="2019",
          subtitle = "")+
  theme(plot.title = element_text(face = "bold", hjust=0.5), plot.subtitle=element_text(hjust=0.5))

plot_grid(p18,p19)


#posterior of difference plots for groups in main effects (supp material)
cowplot::plot_grid(p3, p2, p1, nrow = 2,labels = c("A.", "B.", "C."))

```

### **5. Models of (dis-)assortative mating + relative fitness**

- *Multivariate models that estimate assortative mating (among-pair correlation), response to labile environment (within-pair correlation), and selection gradients for each combination of sex and year.*

``` {r, assortFit, include=T, warning=F, message=F,echo=F}

# Set wd
setwd("C:/Users/nickg/OneDrive/Desktop/R projects/krmp_nest-defense")
require(readr)
#data
data2018<-read.csv("C:/Users/nickg/OneDrive/Desktop/R projects/krmp_nest-defense/data/data_2018.csv")
data2019<-read.csv("C:/Users/nickg/OneDrive/Desktop/R projects/krmp_nest-defense/data/data_2019.csv")

#names(data2018)
data2018$Male<-log(data2018$male_raw+1)*-1
data2018$Female<-log(data2018$female_raw+1)*-1

#names(data2019)
data2019$Male<-log(data2019$male_raw+1)*-1
data2019$Female<-log(data2019$female_raw+1)*-1

#prior for three traits- from Houslay tutorial
prior_E_B_fit_1px = list(R = list(V = diag(c(1,1,0.0001),3,3), nu = 1.002, fix = 3),
                         G = list(G1 = list(V = diag(3), nu = 3,
                                            alpha.mu = rep(0,3),
                                            alpha.V = diag(25^2,3,3))))
#set residual variance for fitness near 0
prior_E_B_fit_1px$R$V[3,3]<-0.0001

```


#### **2018 3 trait model**

```{r, 2018trait, include=T, warning=F, message=T, fig.height=8, fig.width=12}

mod.12 <- MCMCglmm(cbind(Male, Female, rfit) ~ (trait-1),  
                   random = ~us(trait):NestID ,
                   rcov = ~us(trait):units, 
                   family = c("gaussian", "gaussian", "gaussian"),
                   data=data2018, 
                   prior = prior_E_B_fit_1px, 
                   verbose = FALSE,
                   nitt=590000,thin=500,burnin=90000
                   )
summary(mod.12)

plot(mod.12)



#auto-correlation
autocorr.diag(mod.12$VCV)
autocorr(mod.12$Sol)
```

##### **Among and within-pair correlations and selection gradients-- 2018 3 trait model**

``` {r, cor18, include=T, warning=F, message=T}

# posteriors
posteriors_3<-as.mcmc(mod.12$VCV)
posterior.mode(posteriors_3)

# among-pair correlations
pair.correlation_3<-posteriors_3[,"traitFemale:traitMale.NestID"]/
  sqrt(posteriors_3[,"traitFemale:traitFemale.NestID"]*
         posteriors_3[,"traitMale:traitMale.NestID"])

posterior.mode(pair.correlation_3)
HPDinterval(pair.correlation_3)
# within-pair correlations (residuals)
residual.correlation_3<-posteriors_3[,"traitFemale:traitMale.units"]/
  sqrt(posteriors_3[,"traitFemale:traitFemale.units"]*
         posteriors_3[,"traitMale:traitMale.units"])

posterior.mode(residual.correlation_3)
HPDinterval(residual.correlation_3)
#male selection gradient
Male_sel_18<- posteriors_3[,"traitrfit:traitMale.NestID"]/
  (sqrt(posteriors_3[,"traitrfit:traitrfit.NestID"])*
     sqrt(posteriors_3[,"traitMale:traitMale.NestID"]))
posterior.mode(Male_sel_18)
HPDinterval(Male_sel_18)

#female selection gradient
Female_sel_18<- posteriors_3[,"traitrfit:traitFemale.NestID"]/
  (sqrt(posteriors_3[,"traitrfit:traitrfit.NestID"])*
     sqrt(posteriors_3[,"traitFemale:traitFemale.NestID"]))
posterior.mode(Female_sel_18)
HPDinterval(Female_sel_18)
```



#### **2019 3 trait model**

```{r, 2019Assort, include=T, warning=F, message=F, include=T, fig.height=8, fig.width=12}

#prior for three trait- Houslay tutorial 
prior_E_B_fit_1px = list(R = list(V = diag(c(1,1,0.0001),3,3), nu = 1.002, fix = 3),
                         G = list(G1 = list(V = diag(3), nu = 3,
                                            alpha.mu = rep(0,3),
                                            alpha.V = diag(25^2,3,3))))

#model
mod.13 <- MCMCglmm(cbind(Male, Female, rfit) ~ (trait-1),  
                   random = ~us(trait):NestID ,
                   rcov = ~us(trait):units, 
                   family = c("gaussian", "gaussian", "gaussian"),
                   data=data2019, 
                   prior = prior_E_B_fit_1px, 
                   verbose = FALSE,
                   nitt=590000,thin=500,burnin=90000)
summary(mod.13)

plot(mod.13)


#auto-correlation
autocorr.diag(mod.13$VCV)
autocorr(mod.13$Sol)

# posteriors
posteriors1<-as.mcmc(mod.13$VCV)
posterior.mode(posteriors1)

```

##### **Among and within-pair correlations and selection gradients-- 2019 3 trait model**

```{r, cor19, include=T, warning=F, message=F}

# among-pair correlations
pair.correlation_4<-posteriors1[,"traitFemale:traitMale.NestID"]/
  sqrt(posteriors1[,"traitFemale:traitFemale.NestID"]*
         posteriors1[,"traitMale:traitMale.NestID"])

posterior.mode(pair.correlation_4)
HPDinterval(pair.correlation_4)
# within-pair correlations (residuals)
residual.correlation4<-posteriors1[,"traitFemale:traitMale.units"]/
  sqrt(posteriors1[,"traitFemale:traitFemale.units"]*
         posteriors1[,"traitMale:traitMale.units"])

posterior.mode(residual.correlation4)
HPDinterval(residual.correlation4)
#male selection gradient
Male_sel_19<- posteriors1[,"traitrfit:traitMale.NestID"]/
  (sqrt(posteriors1[,"traitrfit:traitrfit.NestID"])*
     sqrt(posteriors1[,"traitMale:traitMale.NestID"]))
posterior.mode(Male_sel_19)
HPDinterval(Male_sel_19)
#female selection gradient
Female_sel_19<- posteriors1[,"traitrfit:traitFemale.NestID"]/
  (sqrt(posteriors1[,"traitrfit:traitrfit.NestID"])*
     sqrt(posteriors1[,"traitFemale:traitFemale.NestID"]))
posterior.mode(Female_sel_19)
HPDinterval(Female_sel_19)
```


##### **Comparing 2018 and 2019**

```{r, compare2018v2019, include=T, warning=F, message=F}
#among
y2018a<-ifelse(pair.correlation_3<0.04291904,1,0)
sum(y2018a)

y2018b<-ifelse(pair.correlation_3>0.04291904,1,0)
sum(y2018b)

y2019a<-ifelse(pair.correlation_4>(-0.334157 ),1,0)
sum(y2019a)

y2019b<-ifelse(pair.correlation_4<(-0.334157 ),1,0)
sum(y2019b)

p<-(sum(y2018a)/1000)*(sum(y2019a)/1000)
p

#within
y2018w<-ifelse(residual.correlation_3>0.04291904,1,0)
sum(y2018w)
y2018x<-ifelse(residual.correlation_3<0.04291904,1,0)
sum(y2018x)
y2019w<-ifelse(residual.correlation4<0.3588893,1,0)
sum(y2019w)
y2019x<-ifelse(residual.correlation4>0.3588893,1,0)
sum(y2019x)

p<-sum(y2018w)/1000*sum(y2019w)/1000
p

##compare sexes within years
m2018<-ifelse(Male_sel_18<0.09374586,1,0)
m2018b<-ifelse(Male_sel_18<0.09374586,0,1)
f2018<-ifelse(Female_sel_18>(-0.4853156),1,0)
f2018b<-ifelse(Female_sel_18>(-0.4853156),0,1)

sum(m2018)
sum(m2018b)
sum(f2018)
sum(f2018b)

p<-sum(m2018)/1000*sum(f2018)/1000
p


m2019<-ifelse(Male_sel_19>-0.2069016,1,0)
m2019b<-ifelse(Male_sel_19>-0.2069016,0,1)
f2019<-ifelse(Female_sel_19<0.08781835,1,0)
f2019b<-ifelse(Female_sel_19<0.08781835,0,1)

sum(m2019)
sum(m2019b)
sum(f2019)
sum(f2019b)

p<-sum(m2019)/1000*sum(f2019)/1000
p

##compare within sex across years
#females
f2018<-ifelse(Female_sel_18>(-0.2069016),1,0)
f2018b<-ifelse(Female_sel_18>(-0.2069016),0,1)
f2019<-ifelse(Female_sel_19<0.09374586,1,0)
f2019b<-ifelse(Female_sel_19<0.09374586,0,1)

sum(f2018)
sum(f2018b)
sum(f2019)
sum(f2019b)

p<-sum(f2018)/1000*sum(f2019)/1000
p

#males
m2018<-ifelse(Male_sel_18<(0.08781835),1,0)
m2018b<-ifelse(Male_sel_18<(0.08781835),0,1)
m2019<-ifelse(Male_sel_19>(-0.4853156),1,0)
m2019b<-ifelse(Male_sel_19>(-0.4853156),0,1)

sum(m2018)
sum(m2018b)
sum(m2019)
sum(m2019b)

p<-sum(m2018)/1000*sum(m2019)/1000
p

```

##### **2018/2019 3 trait model**


``` {r, pooled, nclude=T, warning=F, message=F, include=T, fig.height=8, fig.width=12}
#pooled years----

require(readr)
require(tidyr)
require(MCMCglmm)
#data
data2018<-read.csv("C:/Users/nickg/OneDrive/Desktop/R projects/krmp_nest-defense/data/data_2018.csv")
data2019<-read.csv("C:/Users/nickg/OneDrive/Desktop/R projects/krmp_nest-defense/data/data_2019.csv")



data2018$Male<-log(data2018$male_raw+1)*-1  #log transform 
data2018$Female<-log(data2018$female_raw+1)*-1 #log transform
data2018<-data2018%>%mutate(Site=SiteID)
data2018<-data2018%>% 
  unite(SiteID_Series,c("SiteID", "Year"))
data2018<-data2018%>%mutate(Year="2018")


data2019$Male<-log(data2019$male_raw+1)*-1 #log transform
data2019$Female<-log(data2019$female_raw+1)*-1 #log transform
data2019<-data2019%>%mutate(Site=SiteID)
data2019<-data2019%>% 
  unite(SiteID_Series,c("SiteID", "Year"))
data2019<-data2019%>%mutate(Year="2019")


data_2<-full_join(data2018, data2019) #join datasets

#prior for three traits- from Houslay tutorial
prior_E_B_fit_1px = list(R = list(V = diag(c(1,1,0.0001),3,3), nu = 1.002, fix = 3),
                         G = list(G1 = list(V = diag(3), nu = 3,
                                            alpha.mu = rep(0,3),
                                            alpha.V = diag(25^2,3,3))))
#set residual variance for fitness near 0
prior_E_B_fit_1px$R$V[3,3]<-0.0001

#2018/2019 3 trait model ----
mod.122 <- MCMCglmm(cbind(Male, Female, rfit) ~ (trait-1),  
                   random = ~us(trait):SiteID_Series ,
                   rcov = ~us(trait):units, 
                   family = c("gaussian", "gaussian", "gaussian"),
                   data=data_2, 
                   prior = prior_E_B_fit_1px, 
                   verbose = FALSE,
                   nitt=590000,thin=500,burnin=90000
)

plot(mod.122)
summary(mod.122)

#auto-correlation
autocorr.diag(mod.122$VCV)
autocorr(mod.122$Sol)
```

##### **Among and within-pair correlations and selection gradients-- 2018/2019 3 trait model**

```{r plotmod122, include=T, warning=F, message=F}

# posteriors
require(bayestestR)
require(ggplot2)
posteriors_3<-as.mcmc(mod.122$VCV)


# among-pair correlations
pair.correlation_33<-posteriors_3[,"traitFemale:traitMale.SiteID_Series"]/
  sqrt(posteriors_3[,"traitFemale:traitFemale.SiteID_Series"]*
         posteriors_3[,"traitMale:traitMale.SiteID_Series"])

posterior.mode(pair.correlation_33)
HPDinterval(pair.correlation_33)
paircor<-as.data.frame(pair.correlation_33)
describe_posterior(paircor)
plot(pd(paircor$var1))+theme_classic()

# within-pair correlations (residuals)
residual.correlation_33<-posteriors_3[,"traitFemale:traitMale.units"]/
  sqrt(posteriors_3[,"traitFemale:traitFemale.units"]*
         posteriors_3[,"traitMale:traitMale.units"])

posterior.mode(residual.correlation_33)
HPDinterval(residual.correlation_33)
withincor<-as.data.frame(residual.correlation_33)
describe_posterior(withincor)
plot(pd(withincor$var1))+theme_classic()


#male selection gradient
Male_sel<- posteriors_3[,"traitrfit:traitMale.SiteID_Series"]/
  (sqrt(posteriors_3[,"traitrfit:traitrfit.SiteID_Series"])*
     sqrt(posteriors_3[,"traitMale:traitMale.SiteID_Series"]))
posterior.mode(Male_sel)
HPDinterval(Male_sel)

male<-as.data.frame(Male_sel)
describe_posterior(male)
plot(pd(male$var1))+theme_classic()
#female selection gradient
Female_sel<- posteriors_3[,"traitrfit:traitFemale.SiteID_Series"]/
  (sqrt(posteriors_3[,"traitrfit:traitrfit.SiteID_Series"])*
     sqrt(posteriors_3[,"traitFemale:traitFemale.SiteID_Series"]))
posterior.mode(Female_sel)
HPDinterval(Female_sel)


female<-as.data.frame(Female_sel)
plot(pd(female$var1))+theme_classic()
describe_posterior(female)
plot(pd(female$var1))+theme_classic()
```


##### **Correlation Plot**

```{r corrplot, message=F, warning=F}

### create data frame and within-subject center for plot

require(tidyr)

df_mcmc_cors <- data_frame(Traits = c("Among-pair (2018)",    
                                      "Within-pair (2018)",
                                      "Among-pair (2019)",
                                      "Within-pair (2019)",
                                      "Among-pair (Pooled)",
                                      "Within-pair (Pooled)"
                                        ),
                           Estimate = c(posterior.mode(pair.correlation_3),
                                        posterior.mode(residual.correlation_3),
                                        posterior.mode(pair.correlation_4),
                                        posterior.mode(residual.correlation4),
                                        posterior.mode(pair.correlation_33),
                                        posterior.mode(residual.correlation_33)),
                           Lower = c(HPDinterval(pair.correlation_3)[,"lower"],
                                     HPDinterval(residual.correlation_3)[,"lower"],
                                     HPDinterval(pair.correlation_4)[,"lower"],
                                     HPDinterval(residual.correlation4)[,"lower"],
                                     HPDinterval(pair.correlation_33)[,"lower"],
                                     HPDinterval(residual.correlation_33)[,"lower"]
                                     ),
                           Upper = c(HPDinterval(pair.correlation_3)[,"upper"],
                                     HPDinterval(residual.correlation_3)[,"upper"],
                                     HPDinterval(pair.correlation_4)[,"upper"],
                                     HPDinterval(residual.correlation4)[,"upper"],
                                     HPDinterval(pair.correlation_33)[,"upper"],
                                     HPDinterval(residual.correlation_33)[,"upper"]
                                     ))




correlation1<-ggplot(df_mcmc_cors, aes(x = Traits, y = Estimate)) +
  geom_pointrange(aes(ymin = Lower,
                      ymax = Upper)) +
  geom_hline(yintercept = 0,
             linetype = "dotted",alpha = 0.35, size=0.3) +
  labs(x = "",
       y = "Correlation (Estimate +/- 95% CIs)") +
  ylim(-1,1) +
  coord_flip() + theme_classic()





df_mcmc_cors1 <- data_frame(Traits = c("Female (2018)",    
                                      "Male (2018)",
                                      "Female (2019)",
                                      "Male (2019)",
                                      "Female (Pooled)",
                                      "Male (Pooled)"
),
Estimate = c(posterior.mode(Female_sel_18),
             posterior.mode(Male_sel_18),
             posterior.mode(Female_sel_19),
             posterior.mode(Male_sel_19),
             posterior.mode(Female_sel),
             posterior.mode(Male_sel)
             ),
Lower = c(HPDinterval(Female_sel_18)[,"lower"],
          HPDinterval(Male_sel_18)[,"lower"],
          HPDinterval(Female_sel_19)[,"lower"],
          HPDinterval(Male_sel_19)[,"lower"],
          HPDinterval(Female_sel)[,"lower"],
          HPDinterval(Male_sel)[,"lower"]
),
Upper = c(HPDinterval(Female_sel_18)[,"upper"],
          HPDinterval(Male_sel_18)[,"upper"],
          HPDinterval(Female_sel_19)[,"upper"],
          HPDinterval(Male_sel_19)[,"upper"],
          HPDinterval(Female_sel)[,"upper"],
          HPDinterval(Male_sel)[,"upper"]
          ))




correlation2<-ggplot(df_mcmc_cors1, aes(x = Traits, y = Estimate)) +
  geom_pointrange(aes(ymin = Lower,
                      ymax = Upper)) +
  geom_hline(yintercept = 0,
             linetype = "dotted",alpha = 0.35, size=0.3) +
  labs(x = "",
       y = "Correlation (Estimate +/- 95% CIs)") +
  ylim(-1,1) +
  coord_flip() + theme_classic()

require(cowplot)
plot_grid(correlation1,correlation2, labels = c("A.", "B."), nrow = 1)



```